Nested Sampling with Multiple Objective Functions

### 2018-06-03 doc/ms.tex
Working my way through the introduction still

\section{Multiple objective functions}
In some inference and
statistical mechanics problems, there are two or more scalar functions of
$\xx$ that are relevant. Suppose our prior is $\pi(\xx)$ as before, and
we obtain information that fixes the expected values of two scalar
functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$.
It is well known that the
updated probability distribution that takes into account the constraints is
of the canonical form:
\begin{eqnarray}
p(\xx; \beta_1, \beta_2) &=& \frac{\pi(\xx)\exp\left[\beta_1L_1(\xx)+\beta_2L_2(\xx)\right]}
{Z(\beta_1, \beta_2)}
\end{eqnarray}
where $\beta_1$ and $\beta_2$ are two `inverse temperatures' (the temperatures
themselves are $T_1 = \beta_1^{-1}$ and $T_2 = \beta_2^{-1}$).
In a Bayesian
context, this distribution (when $\beta_1 = \beta_2 = 1$) could be the
posterior distribution for parameters $\xx$
given two datasets, where $L_1$ and $L_2$ are the
likelihood functions for each dataset. In statistical mechanics, $L_1$ might
be $\exp\left(-\textnormal{energy}\right)$ and $L_2$ might be the total
angular momentum, or be a term depending on an external field (so $\beta_2$
would then describe the strength of the external field).

If we are only interested in a single canonical distribution, for example
with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we can estimate its normalising
constant and by running standard Nested Sampling with objective function
$L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
are interested in a range of values for $\beta_1$ and $\beta_2$, and we
want to know the entire partition function $Z(\beta_1, \beta_2)$.
This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions. The algorithm is designed for problems with two objective
functions, but may generalize to problems with more than two (though perhaps
not more than a few).

>>>>

\section{Multiple objective functions}
In some inference and
statistical mechanics problems, there are two or more scalar functions of
$\xx$ that are relevant. Suppose our prior is $\pi(\xx)$ as before, and
we obtain information that fixes the expected values of two scalar
``likelihood'' functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$
(or equivalently, ``energy'' functions $E_1(\xx) = -\ln L_1(\xx)$ and
$E_2 = -\ln L_2(\xx)$). The
updated probability distribution that takes into account the constraints is
of the canonical form \citep{jaynes1957information}:
\begin{align}
p(\xx; \beta_1, \beta_2) &=
    \frac{\pi(\xx)L_1(\xx)^{\beta_1}L_2(\xx)^{\beta_2}}
         {Z(\beta_1, \beta_2)} \\
    &=
    \frac{\pi(\xx)\exp\left[-\beta_1E_1(\xx) - \beta_2E_2(\xx)\right]}
         {Z(\beta_1, \beta_2)}
\end{align}
where $\beta_1$ and $\beta_2$ are two `inverse temperatures' (the temperatures
themselves are $T_1 = \beta_1^{-1}$ and $T_2 = \beta_2^{-1}$).
In a Bayesian
context, this distribution (when $\beta_1 = \beta_2 = 1$) could be the
posterior distribution for parameters $\xx$
given two datasets, where $L_1$ and $L_2$ are the
likelihood functions for each dataset. In statistical mechanics, $E_1$ might
be the total energy and $E_2$ might be the total
angular momentum, or perhaps a term depending on an external field (so $\beta_2$
would then describe the strength of the external field).

If we are only interested in a single canonical distribution, for example
with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we can estimate its normalising
constant and by running standard Nested Sampling with objective function
$L(\xx) = L_1(\xx)^{\beta_1}L_2(\xx)^{\beta_2}$. However, usually we
are interested in a range of values for $\beta_1$ and $\beta_2$, and we
want to know the entire partition function $Z(\beta_1, \beta_2)$.
This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions. The algorithm is presented here for problems with two objective
functions, but the implementation allows for more than two (though the
performance is those cases is likely to be poor).

\
